### 2020-01-03 noise_cutoffs/generate_differing_noise.Rmd
checking differences in noise levels

---
title: "Generate Different Noise Levels"
author: "Robert M Flight"
date: "1/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FTMS.peakCharacterization)
disable_logging()
library(furrr)
future::plan(multicore(workers = 8))
FTMS.peakCharacterization::set_internal_map(furrr::future_map)
```

## Noise??

Based on a previous effort, we are pretty sure we are being *very* conservative in our noise cutoff levels (like 4X too conservative). 
So this document is to create a function that can do the new noise calculation, and then do the full characterization while using different noise cutoffs.

## Plan

Create a custom pipeline for doing the noise cutoff that takes a data file and some different multipliers of the noise cutoff, and then apply it.
We will start with 1, and then go all the way up to 4 in 0.5 increments.
The 4X should be equivalent to what we were doing previously.

## Execute!

### Noise Function

```{r noise_function}
reduce_removing_noise <- function(regions, point_regions, multiplier = 1){
  regions = FTMS.peakCharacterization:::count_overlaps(regions, point_regions)
  nz_counts = regions@elementMetadata$nonzero_counts

  chunk_indices = seq(1, length(nz_counts), by = 10000)

  chunk_perc = purrr::map_dbl(chunk_indices, function(in_index){
    use_density = sliding_density[seq(in_index, min(in_index + 9999, length(sliding_density)))]
    if (max(use_density) > 0) {
      return(stats::quantile(use_density, 0.99))
    } else {
      return(0)
    }
  })

  cutoff_value = median(chunk_perc) * multiplier

  regions = regions[nz_counts > cutoff_value]
  IRanges::reduce(regions)
}
```

### Characterization Function

```{r characterization_function}
characterize_varying_noise = function(raw_file, out_file, multiplier = 1){
  in_char <- CharacterizeMSPeaks$new(raw_file, peak_finder = PeakRegionFinder$new())
  in_char$load_file()
  in_char$filter_raw_scans()
  in_char$zip_ms$peak_finder <- in_char$peak_finder
  in_char$zip_ms$peak_finder$add_data(in_char$zip_ms$raw_ms)
  in_char$zip_ms$raw_ms <- NULL # we drop the raw data to save us some memory
  in_char$zip_ms$peak_finder$sample_id <- in_char$zip_ms$id
  in_char$zip_ms$peak_finder$start_time = Sys.time()
  in_char$zip_ms$peak_finder$progress = TRUE

  in_char$zip_ms$peak_finder$add_regions()
  in_char$zip_ms$peak_finder$peak_regions$peak_regions =
    reduce_removing_noise(in_char$peak_finder$peak_regions$sliding_regions,
                         in_char$peak_finder$peak_regions$frequency_point_regions,
                         multiplier = multiplier)
  in_char$zip_ms$peak_finder$split_peak_regions()
  in_char$zip_ms$peak_finder$remove_double_peaks_in_scans()
  in_char$zip_ms$peak_finder$normalize_data()
  in_char$zip_ms$peak_finder$find_peaks_in_regions()
  in_char$zip_ms$peak_finder$add_offset()
  in_char$zip_ms$peak_finder$sort_ascending_mz()

  in_char$out_file = out_file
  in_char$zip_ms$out_file = in_char$out_file

  in_char$zip_ms$peak_finder$stop_time = Sys.time()
  in_char$summarize()
  in_char$save_peaks()
  in_char$write_zip()
  in_char$zip_ms$cleanup()
}
```

### Run It!

```{r run_noise}
in_file = file.path(here::here(), "data_analysis", "data_input", "161212_unlabeledAAs_1_ECF.mzML")

noise_levels = seq(1, 4, 0.5)

purrr::map(noise_levels, function(in_level){
  out_file = paste0(gsub(".mzML", "", basename(in_file)), "_", in_level, ".zip")
  characterize_varying_noise(in_file, out_file, in_level)
})
```