### 2022-04-13 doc/peakcharacterization_supplemental.Rmd
supplemental stuff and referring to it
closes #22 and #15

---
title: "Supplemental Materials for: Scan-Centric, Frequency-Based Method for Characterizing Peaks from Direct Injection Fourier transform Mass Spectrometry Experiments"
author:
  - Robert M Flight:
      email: robert.flight@uky.edu
      institute: [markey, biochem, rcsirm]
  - Joshua M Mitchell:
      email: jmmi243@uky.edu
      institute: [markey, biochem, rcsirm, ibi]
  - Hunter NB Moseley:
      email: hunter.moseley@uky.edu
      correspondence: true
      institute: [markey, biochem, rcsirm, ibi, tox]
institute:
  - markey: Markey Cancer Center, University of Kentucky, Lexington, KY 40536, USA
  - biochem: Department of Molecular & Cellular Biochemistry, University of Kentucky, Lexington, KY 40536, USA
  - rcsirm: Resource Center for Stable Isotope Resolved Metabolomics, University of Kentucky, Lexington, KY 40536, USA
  - ibi: Institute for Biomedical Informatics, University of Kentucky, Lexington, KY 40536, USA
  - tox: Department of Toxicology and Cancer Biology, University of Kentucky, Lexington, KY 40536, USA
date: "`r Sys.time()`"
output: 
  word_document:
    keep_md: true
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: '`r here::here("doc/peakcharacterization.json")`'
csl: plos-computational-biology.csl
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r get_setup, include = FALSE}
source(here::here("packages.R"))
lapply(list.files(here::here("./R"), full.names = TRUE), source)

figure_count = dn_counter$new("Figure ", prefix2 = "S")
table_count = dn_counter$new("Table ", prefix2 = "S")
equation_count = dn_counter$new("")
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.width = 8, 
                      fig.height = 6, 
                      fig.process = dn_modify_path,
                      dpi = 600,
                      dev.args = list(png = list(type = "cairo")))

format_equation = function(equation_text, label){
  number_out = equation_count$label_text(label)
  paste0('$$\n\\begin{align}\n', equation_text, '&& \\text{(', number_out, ')}', '\n\\end{align}$$')
}
```

## Bruker SolariX ICR Frequency Conversion

In contrast to the m/z to frequency conversion equations used for Thermo-Fisher Fusion and other Orbitrap instruments, the one we have used for Bruker SolariX ICR instruments is simpler:

$$frequency = a + x * \frac{1}{mz} + y * \frac{1}{\sqrt{mz}}$$

## Height - NAP Differences of Corrected vs Raw Scan-Centric

```{r increment_nap_intensity}
figure_count$increment("nap_intensity")
```

Here in `r figure_count$label_text("nap_intensity")` we show the differences in Height - NAP peaks using the corrected and raw scan-centric intensities so they can be compared with the differences we observed between XCalibur and scan-centric peaks.

```{r nap_intensity, dn_id = figure_count, fig.width = 16, fig.height = 8}
nap_fontsize = 12
tar_load(aa_diffs_filtersd_1ecf)
tar_load(aa_diffs_filtersd_2ecf)

all_diffs = rbind(aa_diffs_filtersd_1ecf$all_diffs %>%
                    dplyr::mutate(sample = "1ecf"),
                  aa_diffs_filtersd_2ecf$all_diffs %>%
                    dplyr::mutate(sample = "2ecf"))

aa_diffplot = all_diffs %>%
  dplyr::filter(source %in% "char_raw_corrected") %>%
  dplyr::mutate(diff = -1 * diff) %>%
  ggplot(aes(x = n_missing, y = diff)) +
  geom_point() +
  labs(x = "Number of Missing Scans",
         y = "Corrected - Raw Scan-Centric NAP - Height Differences") +
  theme(axis.title.y = element_text(size = nap_fontsize))

aa_diffplot
```

`r figure_count$label_text("nap_intensity")`.
The peak-peak NAP - intensity log differences between corrected and raw scan-centric peak heights as a function of the number of scans the peaks were not found in across all of the amino acid assignments in EMFs with more than a single peak in both ECF samples.

## ICI-Kt Median Correlation Outliers

```{r increment_icikt}
figure_count$increment("icikt_outlier")
```

For the NSCLC dataset, we used information-content-informed Kendall-tau (ICI-Kt) correlation within each of non-cancer and cancer sample groups to determine outliers that should be removed prior to statistical testing.
In `r figure_count$label_text("icikt_outlier")` we show the distribution of median ICI-Kt values and the samples determined as outliers.

```{r icikt_outlier, dn_id = figure_count}
tar_load(qcqa)
outliers = qcqa$outlier_data

outliers %>%
  ggplot(aes(x = sample_class, y = med_cor, color = outlier, group = sample_class)) +
  geom_sina() +
  theme(legend.position = c(0.1, 0.9)) +
  labs(x = "Sample Class", y = "Median ICI-Kt")
```

`r figure_count$label_text("icikt_outlier")`.
Sina plot of median ICI-Kt correlation for samples within each of the cancer and non-cancer sample groups with outlier status indicated for each sample.

## Incorrect Normalization in P-Value Differences

```{r increment_p_value}
figure_count$increment("p_value")
```

During the analysis of the p-values between the two classes of NSCLC samples, we didn't originally have the median intensities for the corrected peaks.
Therefore, we went ahead and normalized the corrected intensities in each sample with the raw median intensities in each sample, and compared the p-values from each method.
`r figure_count$label_text("p_value")` shows how this incorrect median normalization affects the difference in p-values between corrected and raw intensity.

```{r p_value, dn_id = figure_count}
lung_compare = tar_read(lung_compare_wrong)
lung_compare = lung_compare %>%
  dplyr::mutate(compare_id = dplyr::case_when(
    id.test %in% "corrected" ~ "Corrected",
    id.test %in% "xcalibur" ~ "XCalibur",
    id.test %in% "msnbase" ~ "MSnbase"
  ))

lung_compare_corrected = lung_compare %>%
  dplyr::filter(compare_id %in% "Corrected")

lung_compare_ttest = tar_read(lung_wrong_ttest)

long_ttest = lung_compare_ttest %>%
  dplyr::select(p.adjust, comparison, estimate, conf.low, conf.high) %>%
  tidyr::pivot_longer(cols = !comparison, names_to = "measure", values_to = "value") %>%
  dplyr::mutate(compare_id = dplyr::case_when(
    comparison %in% "corrected" ~ "Corrected",
    comparison %in% "xcalibur" ~ "XCalibur",
    comparison %in% "msnbase" ~ "MSnbase"
  )) %>%
  dplyr::filter(compare_id %in% "Corrected")

long_pvalue = long_ttest %>%
  dplyr::filter(measure %in% "p.adjust") %>%
  dplyr::filter(compare_id %in% "Corrected")

long_measure = long_ttest %>%
  dplyr::filter(!(measure %in% "p.adjust")) %>%
  dplyr::filter(compare_id %in% "Corrected")

compare_sina = lung_compare_corrected %>%
  ggplot(aes(x = compare_id, y = p_diff)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_sina(alpha = 0.5) +
  geom_point(data = long_measure, aes(x = compare_id, y = value), color = "green") +
  geom_text(data = long_pvalue, aes(x = compare_id, y = -1, label = format(value, digits = 2))) +
  labs(x = "", y = "Log-P-Value Difference")

compare_pair = lung_compare_corrected %>%
  ggplot(aes(x = log_p.ref, y = log_p.test)) +
  geom_abline(slope = 1, color = "red") +
  geom_point(alpha = 0.5) +
  labs(x = "Log-P-Value Raw",
         y = "Log-P-Value Corrected")
compare_pair | compare_sina + plot_annotation(tag_levels = "A")
```

`r figure_count$label_text("p_value")`.
**A**: Log-p-values generated by comparing non-cancer and cancer sample IMFs using peak intensities from raw and corrected peak intensities, where the corrected intensities were normalized using the sample raw median intensities.
Red line denotes perfect agreement.
**D**: Sina plot of differences in the log-p-values generated by corrected intensity normalized by the raw median intensities.
Also shown are the Bonferroni adjusted p-values from a t-test of the log-p-value differences for each method.
Green points denote the high, mean, and low-confidence limits reported from the t-test.